---
title: "unemployment-rate-by-president"
author: "Justin Shotwell"
date: "4/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE}

library(tidyverse)
library(lubridate)
library(ggthemes)
library(CGPfunctions)

theme_set(theme_tufte())

```

The csv file that I am using is one that I put together myself. The data is from the bureau of labor services(bls.gov). I added the dates of each president and the president's political party before reading into R. 

As the inauguration is on the 20th of the month, I credited the outgoing president with that month. For examples of assassination or impeachment, I gave credit for the month in question to the president that was president for more than half the month.

For example, January of 2017 would be creditied to Barack Obama because he was president for 19 days and Donald Trump was president for 12 days.

```{r, message=FALSE}

unemployment_raw <- read_csv("unemployment rate history - Sheet1.csv") %>%
  mutate(date = mdy(date))

str(unemployment_raw)

```

```{r}

by_president <- unemployment_raw %>%
  group_by(president) %>%
  summarize(start = value[1],
            end = value[length(value)],
            party = party[1]) %>%
  gather(attribute, value, -president, -party) %>%
  mutate(attribute = as_factor(attribute))

```

```{r, warning=FALSE}

custom_colors <- if_else(by_president$party == "Democrat", "blue", "red")

by_president %>%
  newggslopegraph(attribute,
                  value,
                  president,
                  Title = "Unemployment Rate Change by President", 
                  SubTitle = "Does each President leave office with a better or worse rate?", 
                  Caption = NULL,
                  YTextSize = 4,
                  DataTextSize = 3.5,
                  LineColor = custom_colors,
                  ThemeChoice = "tufte")

```

